KidsAIR: Main analysis

### 2020-12-17 main_analysis_final.Rmd
analysis updates - splines

---
title: 'KidsAIR: Main analysis'
author: "Ethan Walker"
date: "Updated 14 Dec 2020"
output: pdf_document
---

@@

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_total = as.factor(lrti_events_di_total),
         person_time_at_risk_iqr = person_time_at_risk/20) 
  #filter(gender == "Female")
  #filter(home_sqft_2level == "1500+")

analysis_data_outliers <- analysis_data %>% 
  #rownames_to_column() %>% 
  #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
  mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk),
         lrti_events_total = as.factor(lrti_events_total),
         lrti_events_di_total = as.factor(lrti_events_di_total))

analysis_data_winter1 <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_winter = as.factor(lrti_events_di_winter)) %>% 
  filter(person_time_at_risk_winter > 0) %>% 
  filter(winter_id == 1)

analysis_data_winter2 <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_winter = as.factor(lrti_events_di_winter)) %>% 
  filter(person_time_at_risk_winter > 0) %>% 
  filter(winter_id == 2)

analysis_data_noak <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_total = as.factor(lrti_events_di_total),
         person_time_at_risk_iqr = person_time_at_risk/20) %>% 
  filter(area != "AK")
```

@@

# Run primary model and print results
itt_results_int <<- glmer(lrti_events_di_total ~ treatment_assigned*int_var + 
                          child_age + 
                          person_time_at_risk + (1 | cohort:area),
                          data = new_model_data, family = binomial, nAGQ = 10,
                          control = glmerControl(optimizer = "bobyqa"))
#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
#summary(itt_pm_results_int)

@@

@@

int_function(analysis_data, "residents_smoke")
n_table
save_contrasts
pm_summary_table
save_contrasts_full

@@

tidy_results <- tidy(itt_filter_type_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_filter_type", "", term),
         model = "Filter type") %>% 
  mutate(estimate = round(exp(estimate), digits = 2),
         conf.low = round(exp(conf.low), digits = 2),
         conf.high = round(exp(conf.high), digits = 2)) %>% 
  dplyr::select(model, term, estimate, conf.low, conf.high)
tidy_results

table(model.frame(itt_filter_type_results)$treatment_filter_type)

@@

# Run primary model and print results
itt_pm_results_int <<- lmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
                              child_age + 
                       (1 | cohort:area), data = new_model_data)
#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
#summary(itt_pm_results_int)

@@

@@

int_function(analysis_data_winter1, "burn_level_3level")
n_table
save_contrasts
pm_summary_table
save_contrasts_full

@@

# Run primary model and print results
er_results_int <<- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*int_var + 
                          child_age + 
                          education_3level + 
                          residents_smoke + 
                          person_time_at_risk_winter +
                          (1 | cohort:area), 
                          data = new_model_data, family = binomial, nAGQ = 10,
                          control = glmerControl(optimizer = "bobyqa"))
#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
#summary(itt_pm_results_int)

@@

@@

int_function(analysis_data_winter1, "gender")
save_contrasts
pm_summary_table

@@

# Check PM concentrations for subgroups
pm_summary_table <- analysis_data_winter1 %>% 
  filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
           !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
  #group_by(treatment_assigned) %>% 
  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
pm_summary_table
```

```{rr, eval=FALSE, include=FALSE}
# ER framework using a Spline for PM

@@

model_pm_spline <- gamm4(lrti_events_di_winter ~ s(pm_mean_sampling_period_iqr, 
                          fx = TRUE, bs = "tp", k = 6) + 
                          child_age +  education_3level + residents_smoke +
                          person_time_at_risk_winter, 
                          data = analysis_data_winter1, family = binomial(link = "logit"), 
                          random = ~(1 | home:cohort:area))
plot(model_pm_spline$gam, seWithMean = TRUE, select = 1, shade = TRUE,
     xlab = expression(paste("IQR increase in area PM"[2.5], 
                             " (", mu, g/m^3, ")")),
     ylab = expression(paste("OR, spline trend for area PM  "[2.5])))
summary(model_pm_spline$gam)
summary(model_pm_spline$mer)

# Not sure how to add the offset in
er_results <- glmer.nb(lrti_events_winter ~ pm_mean_sampling_period_iqr + child_age + 
                          education_3level + residents_smoke + 
                          (1 | home:cohort:area) + offset(log_person_time_at_risk_winter), 
                          data = analysis_data_winter1)
```

>>>>

---
title: 'KidsAIR: Main analysis'
author: "Ethan Walker"
date: "Updated 16 Dec 2020"
output: pdf_document
---

@@

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_total = as.factor(lrti_events_di_total),
         person_time_at_risk_iqr = person_time_at_risk/20,
         child_age = factor(child_age, levels = c("< 1", "1+"))) 
  #filter(gender == "Female")
  #filter(home_sqft_2level == "1500+")

analysis_data_outliers <- analysis_data %>% 
  #rownames_to_column() %>% 
  #mutate(filter_outliers = if_else(rowname %in% c(57), 1, 0)) %>% 
  mutate(log_person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 2.33, log_person_time_at_risk)) %>% 
  mutate(person_time_at_risk = if_else(area == "AK" & child_id_num == 31, 72, person_time_at_risk),
         lrti_events_total = as.factor(lrti_events_total),
         lrti_events_di_total = as.factor(lrti_events_di_total),
         child_age = factor(child_age, levels = c("< 1", "1+")))

analysis_data_winter1 <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_winter = as.factor(lrti_events_di_winter),
         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
  filter(person_time_at_risk_winter > 0) %>% 
  filter(winter_id == 1)

analysis_data_winter2 <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_2obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_winter = as.factor(lrti_events_di_winter),
         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
  filter(person_time_at_risk_winter > 0) %>% 
  filter(winter_id == 2)

analysis_data_noak <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  mutate(pm_mean_sampling_period_iqr = pm_mean_sampling_period/25,
         pm_at_home_sampling_period_iqr = pm_at_home_sampling_period/25,
         lrti_events_di_total = as.factor(lrti_events_di_total),
         person_time_at_risk_iqr = person_time_at_risk/20,
         child_age = factor(child_age, levels = c("< 1", "1+"))) %>% 
  filter(area != "AK")
```

@@

# Run primary model and print results
itt_results_int <<- glmer(lrti_events_di_total ~ treatment_assigned*int_var + 
                          child_age + 
                          person_time_at_risk + (1 | cohort:area),
                          data = new_model_data, family = binomial, nAGQ = 10,
                          control = glmerControl(optimizer = "bobyqa"))
#write_rds(itt_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
save_summary <<- summary(itt_results_int)

@@

lrti_summary_table <<- new_model_data %>% 
  filter(!is.na(int_var)) %>% 
  group_by(int_var, treatment_assigned) %>% 
  mutate(lrti_events_num = as.numeric(lrti_events_di_total),
         lrti_events_num = lrti_events_num - 1) %>%  
  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
            "Group %" = `Total LRTI`/n()*100)

@@

int_function(analysis_data, "residents_smoke")
n_table
save_contrasts
pm_summary_table
lrti_summary_table
save_summary
save_contrasts_full

@@

tidy_results <- tidy(itt_winter1_results, conf.int = TRUE) %>% 
  mutate(group_filter = if_else(grepl("treatment", term), 1, 0)) %>% 
  filter(group_filter == 1) %>% 
  mutate(term = gsub("treatment_filter_type", "", term),
         model = "Filter type") %>% 
  mutate(estimate = round(exp(estimate), digits = 2),
         conf.low = round(exp(conf.low), digits = 2),
         conf.high = round(exp(conf.high), digits = 2)) %>% 
  dplyr::select(model, term, estimate, conf.low, conf.high)
tidy_results

table(model.frame(itt_winter1_results)$treatment_assigned)

@@

# Run primary model and print results
itt_pm_results_int <<- lmer(log(pm_mean_sampling_period) ~ treatment_assigned*int_var + 
                              child_age + 
                       (1 | cohort:area), data = new_model_data)
#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
save_summary <<- summary(itt_pm_results_int)

@@

lrti_summary_table <<- new_model_data %>% 
  filter(!is.na(int_var)) %>% 
  group_by(int_var, treatment_assigned) %>% 
  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
         lrti_events_num = lrti_events_num - 1) %>%  
  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
            "Group %" = `Total LRTI`/n()*100)

@@

int_function(analysis_data_winter1, "burn_level_3level")
n_table
save_contrasts
pm_summary_table
lrti_summary_table
save_summary
save_contrasts_full

@@

# Run primary model and print results
er_results_int <<- glmer(lrti_events_di_winter ~ pm_mean_sampling_period_iqr*int_var + 
                          child_age + 
                          education_3level + 
                          #residents_smoke + 
                          person_time_at_risk_winter +
                          (1 | cohort:area), 
                          data = new_model_data, family = binomial, nAGQ = 10,
                          control = glmerControl(optimizer = "bobyqa"))
#write_rds(itt_pm_results_int, paste0(file_path, "Output/itt_pm_results_int_age.rds"))
save_summary <<- summary(er_results_int)

@@

lrti_summary_table <<- new_model_data %>% 
  filter(!is.na(int_var)) %>% 
  group_by(int_var) %>% 
  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
         lrti_events_num = lrti_events_num - 1) %>%  
  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
            "Group %" = `Total LRTI`/n()*100)

@@

int_function(analysis_data_winter1, "residents_smoke")
save_contrasts
pm_summary_table
lrti_summary_table
save_summary

@@

# Check PM concentrations for subgroups
pm_summary_table <- analysis_data_winter1 %>% 
  filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
           !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
  #group_by(treatment_assigned) %>% 
  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
pm_summary_table

pm_summary_table <- analysis_data_winter1 %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  group_by(treatment_filter_type) %>% 
  summarize("Mean PM" = mean(pm_mean_sampling_period, na.rm = TRUE),
            "n" = n(),
            "SD PM" = sd(pm_mean_sampling_period, na.rm = TRUE),
            "Min PM" = min(pm_mean_sampling_period, na.rm = TRUE),
            "Median PM" = median(pm_mean_sampling_period, na.rm = TRUE),
            "Max PM" = max(pm_mean_sampling_period, na.rm = TRUE))
pm_summary_table

lrti_summary_table <- analysis_data_winter1 %>% 
  filter(!is.na(lrti_events_di_winter) & !is.na(pm_mean_sampling_period) &
           !is.na(child_age) & !is.na(education_3level) & !is.na(residents_smoke)) %>% 
  #group_by(treatment_assigned) %>% 
  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
         lrti_events_num = lrti_events_num - 1) %>%  
  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
            "Group %" = `Total LRTI`/n()*100)
lrti_summary_table

lrti_summary_table <- analysis_data_winter1 %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  mutate(lrti_events_num = as.numeric(lrti_events_di_winter),
         lrti_events_num = lrti_events_num - 1) %>% 
  group_by(treatment_filter_type) %>% 
  summarize("Total LRTI" = sum(lrti_events_num, na.rm = TRUE),
            "Group %" = `Total LRTI`/n()*100)
lrti_summary_table
```

```{r, eval=FALSE, include=FALSE}
# ER framework using a Spline for PM

spline_data <- analysis_data_winter1 %>% 
  filter(pm_mean_sampling_period < 200) %>% 
  mutate(pm_cat = cut(pm_mean_sampling_period,
                      breaks = c(0, 10, 20, 35, 300),
                      labels = c("0 to 10", "10 to 20",
                                 "20 to 35", "35+")))

@@

model_pm_spline <- gamm4(lrti_events_di_winter ~ s(pm_mean_sampling_period, 
                          fx = TRUE, bs = "tp", k = 5) + 
                          child_age +  education_3level + residents_smoke +
                          person_time_at_risk_winter, 
                          data = spline_data, family = binomial(link = "logit"), 
                          random = ~(1 | cohort:area))
plot(model_pm_spline$gam, seWithMean = TRUE, select = 1, shade = TRUE,
     xlab = expression(paste("Area PM"[2.5], 
                             " (", mu, g/m^3, ")")),
     ylab = expression(paste("OR, spline trend for area PM  "[2.5])))
summary(model_pm_spline$gam)
summary(model_pm_spline$mer)

# Try using PM grouped by quartiles

er_results <- glmer(lrti_events_di_winter ~ pm_cat + child_age + 
                          education_3level + residents_smoke + person_time_at_risk_winter +
                          (1 | cohort:area), 
                          data = spline_data, family = binomial, nAGQ = 10,
                          control = glmerControl(optimizer = "bobyqa"))
summary(er_results)
```
